---
title: "Homework 6"
author: "Emma Staut"
date: "3/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Team members: Emma Staut
Repository link: https://github.com/estaut/Homework-6.git

# Part 1: Data Cleaning

1. Data cleaning: replace the values "IAP", "DK", and "NA" with the value NA.
```{r, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)

HAPPY <- readRDS("~/Homework-6/data/HAPPY.rds")

HAPPY <- replace(HAPPY, HAPPY == "IAP", NA)
HAPPY <- replace(HAPPY, HAPPY == "DK", NA)
HAPPY <- replace(HAPPY, HAPPY == "NA", NA)
head(HAPPY)
```

2. Check the type of variable and cast into correct type. For age, change "89 OR OLDER" to 89 and cast to numeric.
3. Bring all levels of factors to a sensible order.
```{r}
str(HAPPY)
HAPPY1 =
  HAPPY %>% mutate(AGE = replace(AGE, AGE == '89 OR OLDER', 89)) %>%
  mutate(AGE = as.numeric(AGE),
         HAPPY = as.factor(HAPPY),
         SEX = as.factor(SEX), 
         MARITAL = as.factor(MARITAL),
         DEGREE = as.factor(DEGREE),
         FINRELA = as.factor(FINRELA),
         HEALTH = as.factor(HEALTH),
         PARTYID = as.factor(PARTYID),
         POLVIEWS = as.factor(POLVIEWS)) %>% 
  mutate(MARITAL = factor(MARITAL, levels = c("NEVER MARRIED",
                                              "MARRIED",
                                              "SEPARATED",
                                              "DIVORCED",
                                              "WIDOWED"))) %>%
  mutate(DEGREE = factor(DEGREE, levels = c("LT HIGH SCHOOL",
                                                     "HIGH SCHOOL",
                                                     "JUNIOR COLLEGE",
                                                     "BACHELOR",
                                                     "GRADUATE"))) %>%
  mutate(FINRELA = factor(FINRELA, levels = c("FAR BELOW AVERAGE",
                                              "BELOW AVERAGE",
                                              "AVERAGE",
                                              "ABOVE AVERAGE",
                                              "FAR ABOVE AVERAGE"))) %>%
  mutate(HEALTH = factor(HEALTH, levels = c("POOR",
                                            "FAIR",
                                            "GOOD",
                                            "EXCELLENT"))) %>%
  mutate(PARTYID = factor(PARTYID, levels = c("OTHER PARTY",
                                              "STRONG REPUBLICAN",
                                              "NOT STR REPUBLICAN",
                                              "IND,NEAR REP",
                                              "INDEPENDENT",
                                              "IND,NEAR DEM",
                                              "NOT STR DEMOCRAT",
                                              "STRONG DEMOCRAT"))) %>%
  mutate(POLVIEWS = factor(POLVIEWS, levels = c("EXTRMLY CONSERVATIVE",
                                                "CONSERVATIVE",
                                                "SLGHTLY CONSERVATIVE",
                                                "MODERATE",
                                                "SLIGHTLY LIBERAL",
                                                "LIBERAL",
                                                "EXTREMELY LIBERAL")))

str(HAPPY1)
```
Save the final data set as an RDS file using "saveRDS".
```{r}
saveRDS(HAPPY1, file='happy1.rds')
```

# Part 2: Individual Exploration

I will explore the relationship between happiness and health between the two sexes.
```{r, warning=FALSE}
library(ggmosaic)
HAPPY1 %>% ggplot() + geom_mosaic(aes(x=product(SEX), fill=HAPPY, weight=1)) + facet_grid(~HEALTH)
```
As we can see from the graph, the relationship between happiness and health is the same between the sexes at each health level. However, happiness does appear to depend on health. The graph shows that the healthier a person is, the happier they are. This is shown by the increasing number of "Very happy" responses as health increases, and the decreasing number of "Not so happy" responses as health increases. Approximately the same number of people responded "Pretty happy", regardless of health level.